---
title: "Riddler - 2017-08-25"
output: html_notebook
---

## Riddler classic

*What's the minimum size of tetrahedron that can hold four spheres each with radius 1?*

First step is to assume that the problem is symmetric w.r.t. the four sides of the tetrahedron, so that we can consider a representative side in two dimensions and then generalise the answer to three.

This is important because: (i) I have serious doubts about my laptop's ability to converge to a good solution in three dimensions, and (ii) because I'd like to avoid wrestling with 3D graphics if at all possible.

The code can be broken down into a few sections:

1. Generating starting positions for the circles and triangles

2. Creating a few helper functions to calculate area and check for collisions

3. The core annealing algorithm

4. Generating GIFs from the results

# Setup

```{r setup}
library(ggplot2)
library(ggthemes)
library(tibble)
library(magrittr)
library(geometry)
library(animation)

saveGIF(plot(1:3)) # seem to need to call the function
# once unsuccessfully before I can fix the file path below
ani.options(convert = shortPathName(ani.options("convert")))

```

```{r}
# arbitrary start points for each circle
centers <- tribble(
    ~x,     ~y,
  -2.5,    2.5,
     1,      1,
     0,   -3.5
)

```

```{r}
# 100 points around each circle for edge detection
gen_unit_circle <- function(n_pts) {
  r <- 1:n_pts / n_pts * 2 * pi
  data.frame(
    x = cos(r),
    y = sin(r)
  )
}

gen_circle_pts <- function(unit_c, cen) {
  L <- apply(cen, 1, function(i) {
    data.frame(
      x = unit_c$x + i[1],
      y = unit_c$y + i[2]
    )
    })
  do.call(rbind, L)
}

unit_c <- gen_unit_circle(100)
c_pts <- gen_circle_pts(unit_c, centers)

```

```{r}
# arbitrary start points for triangle
tri <- tribble(
   ~x,  ~y,
    0,  15,
  -15, -8,
   15, -8
)

```

# Functions

```{r}
# area of a cartesian triangle
area <- function(p) {
  mat <- array(c(p$x, p$y, rep(1,3)), dim = c(3,3))
  abs(det(mat) / 2)
}

```

```{r}
# check for edge violations (circles outside triangle)
check_edges <- function(c, p) {
  c_rows <- dim(c)[1]
  p_rows <- dim(p)[1]
  pts <- p
  pts[(p_rows+1):(p_rows+c_rows),] <- c
  hull <- convhulln(pts)
  return(max(hull) <= p_rows)
}

```

```{r}
# check for circles touching
ball_collide <- function(cen) min(dist(cen)) <= 2

```

# Annealing

```{r}
sim_anneal <- function(cen, c_pts, tri, rounds) {
  # browser()
  # PARAMETERS
  start_temp <- 4
  temp <- start_temp
  decay <- (1 - (1 / (rounds/11)))
  unit_c <- gen_unit_circle(100)
  # INITIALISE HISTORY
  history <- tri
  history$c_x <- cen$x
  history$c_y <- cen$y
  history$t <- 0L
  history$area <- area(tri)
  history$temp <- start_temp
  history$check <- TRUE
  history$prb_keep <- 1
  history$keep <- TRUE
    for (t in 1:rounds) {
    old_area <- area(tri)
    # STEP ONE: NEW TRIANGLE
    new_tri <- tri + rnorm(6)*temp/start_temp
    ## probability of keeping new tri
    new_area <- area(new_tri)
    if (new_area < old_area) {
      prb_keep <- 1
    } else {
      prb_keep <- exp((old_area - new_area) / temp)
    }
    check <- check_edges(c_pts, new_tri)
    if (!check) prb_keep <- 0
    ## keep/discard tri
    if (runif(1) < prb_keep) {
      tri <- new_tri
      keep <- TRUE
    } else {keep <- FALSE}
    # STEP TWO: NEW CIRCLES
    new_cen <- cen + rnorm(6)*temp/start_temp
    ## decide whether to keep new circles
    if (ball_collide(new_cen)) {
      cen_keep <- FALSE
    } else {
      new_c_pts <- gen_circle_pts(unit_c, new_cen)
      if (!check_edges(new_c_pts, tri)) {
        cen_keep <- FALSE
      } else {
        cen_keep <- TRUE
      }
    }
    if (cen_keep) {
      cen <- new_cen
      c_pts <- new_c_pts
    }
    # RECORD KEEPING
    new_history <- new_tri
    new_history$c_x <- cen$x
    new_history$c_y <- cen$y
    new_history$t <- t
    new_history$area <- new_area
    new_history$temp <- temp
    new_history$check <- check
    new_history$prb_keep <- prb_keep
    new_history$keep <- keep
    history <- rbind(history, new_history)
    temp <- temp * decay # exponential
    if (t %% 200 == 0) cat(paste0(t,"."))
  }
  return(history)
}

res <- sim_anneal(centers, c_pts, tri, 20000)

print(min(res$area[res$keep]))

```

# Plotting

```{r}
circle_plot <- function(res_data, t_val) {
  # browser()
  res_data <- res_data[res_data$t == t_val,]
  chk <- max(res_data$check)
  tri_color <- ifelse(chk,"navy","red")
  tri_fill <- ifelse(chk,"navy",NA)
  g <- ggplot(res_data) +
    geom_point(aes(c_x,c_y),
               shape = 21, size = 18, fill = "green") + 
    geom_polygon(aes(x,y),
                 color = tri_color, fill = tri_fill, alpha = 0.1) +
    coord_fixed(ratio = 1, xlim = c(-11,11), ylim = c(-10,11)) +
    theme_solid() + theme(legend.position = "none")
  print(g)
}

```

```{r}
saveGIF(for (t in seq(0,20000,200)) {
  circle_plot(res, t)
}, interval = 0.1)

```

```{r}
nice_plot <- function(res_data, t_val) {
  # browser()
  res_data <- res_data[res_data$t == t_val,]
  chk <- max(res_data$t) != max_t # define
  tri_color <- ifelse(chk,"blue3","#33cc33")
  tri_fill <- ifelse(chk, "blue3" ,"#33cc33")
  g <- ggplot(res_data) +
    geom_polygon(aes(x,y),
                 color = tri_color, fill = tri_fill, alpha = 0.1) +
    geom_text(data = data.frame(x = -9.3, y = -11.2),
              aes(x, y, label = "Area: "),
              hjust = 1,
              size = 5) + 
    geom_text(aes(x = -9.3, y = -11.2, label = round(area,2)),
              hjust = 0, color = tri_color,
              size = 5) +
    geom_text(data = data.frame(x = -9.3, y = -12.2),
              aes(x, y, label = "Temp: "),
              hjust = 1,
              size = 5) + 
    geom_text(aes(x = -9.3, y = -12.2, label = round(temp,2)),
              hjust = 0, color = tri_color,
              size = 5) +
    geom_text(data = data.frame(x = -9.3, y = -13.2),
              aes(x, y, label = "Round: "),
              hjust = 1,
              size = 5) + 
    geom_text(aes(x = -9.3, y = -13.2, label = round(t,2)),
              hjust = 0, color = tri_color,
              size = 5) +
    geom_point(aes(c_x,c_y),
               shape = 21, size = 18, 
               fill = "lightpink") + 
    coord_fixed(ratio = 1, xlim = c(-11,11), ylim = c(-13,8)) +
    theme_solid() + 
    theme(legend.position = "none",
          panel.background = element_rect(fill = "grey85"))
  print(g)
}

```

```{r}
res2 <- res[res$keep,]
t_vals <- sort(unique(res2$t))
t_vals <- t_vals[seq(0,length(t_vals),length(t_vals)/40)]
t_vals <- t_vals[t_vals > 4000]
t_vals <- c(rep(0L,22), seq(1,600,2), seq(605,4000,8), 
            t_vals, rep(max(t_vals),14))
t_vals <- t_vals[t_vals %in% res$t[res$keep]]
max_t <- max(t_vals)

saveGIF(for (t in t_vals) {
  nice_plot(res, t)
}, interval = 0.08)

```

```{r}
time_string <- function() gsub("[ :]", "_", Sys.time())

saveRDS(res, paste0("res_",time_string(),".RDS"))

```






